A load balancing scheme for MMOGs

### 2009-03-26 trunk/textos/dissertation/MTAP_LaTeX/pub/tex/latex/svjour3/global/balancing.tex

\newcommand{\ggp}{ProGReGA}
%\newcommand{\ggp}{PGRGA}
\newcommand{\ggpmeaning}{proportional greedy region growing algorithm}
\newcommand{\ggpk}{\mbox{\ggp-KH}}
\newcommand{\ggpkmeaning}{proportional greedy region growing algorithm keeping heaviest cell}
\newcommand{\ggpf}{\ggp-KF}
\newcommand{\ggpfmeaning}{proportional greedy region growing algorithm keeping usage fraction}
\newcommand{\bfa}{BFBCT}
\newcommand{\bfameaning}{best-fit based cell transference}
\newcommand{\kl}{Kernighan\-Lin}
\newcommand{\klalg}{Kernighan-Lin}
\newcommand{\wtodiv}{weight\_to\_divide}
\newcommand{\freecap}{free\_capacity}
\newcommand{\rlist}{region\_list}
\newcommand{\clist}{cell\_list}
\newcommand{\locgroup}{local\_group}
\newcommand{\avguse}{average\_usage}
\newcommand{\wlocal}{local\_weight}
\newcommand{\caplocal}{local\_capacity}
%\newcommand{\wshare}{por\text{\emph{\c{c}}}\tilde{a}o\_da\_carga}
\newcommand{\wshare}{weight\_share}
\newcommand{\wlose}{weight\_to\_lose}
\newcommand{\destreg}{destination\_regions}
\newcommand{\regcap}{free\_capacity}
\newcommand{\wtoreg}{weight\_to\_this\_region}

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2006/03/15
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
gsave
newpath
  20 20 moveto
  20 220 lineto
  220 220 lineto
  220 20 lineto
closepath
2 setlinewidth
gsave
  .4 setgray fill
grestore
stroke
grestore
\end{filecontents*}
%
\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallextended]{svjour3}     % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}         % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage[latin1]{inputenc}
\usepackage{algorithmic}
\usepackage{algorithm}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}
%[TODO:ajustar TODAS as \refs]
\title{A load balancing scheme for MMOGs %\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

Normally, players can freely move their avatars throughout the game world. This makes possible the formation of \emph{hotspots} \cite{ahmed2008mol}, around which the players are more concentrated than in other regions of the virtual environment. Moreover, many massively multiplayer online RPGs not only permit but also stimulate, to some extent, the formation of these points of interest. In the worlds of these MMORPGs, there are entire cities, where the players meet to chat, exchange virtual goods or even fight, and there are also desertic areas, with few attractions for the players, and where the number of avatars is relatively small compared to other places of the environment.

@@

The servers receive the action performed by a player, calculate its outcome and send it to all interested players, who are usually those whose avatars are close to the avatar of the first player. If two players are split between different servers, each of these need not only to send the state update to the player served by it, but it has also to send the update to the server to which the other player is connected. This server, in turn, forwards that state update to the other player.

%[TODO:por figura do envio de estados através de servidores diferentes. duas figuras: no mesmo servidor, e através de diferentes servidores.]

@@

Para exemplificar o funcionamento do algoritmo, pode-se observar a Figura \ref{fig:lee2}. Todos os servidores têm a mesma capacidade, podendo cada um comportar 100 usuários. Primeiro, o servidor iniciador, $S_6$, é inserido em SELECIONADOS e seus vizinhos ($S_2$, $S_5$, $S_7$ e  $S_{10}$) são adicionados a CANDIDATOS (Figura \ref{fig:lee2}(a)). Então, $S_7$, que tem a menor carga de trabalho dentre os servidore em CANDIDATOS, é selecionado e convidado a participar da distribuição de carga. Quando $S_7$ responde a $S_6$ com a informação de carga de seus vizinhos ($S_3$, $S_6$, $S_8$ e $S_{11}$), $S_7$ é inserido em SELECIONADOS e seus vizinhos, exceto $S_6$, são adicionados a CANDIDATOS (Figura \ref{fig:lee2}(b)). Agora, $S_{11}$, que tem a menor carga de trabalho dentre os servidores em CANDIDATOS, é selecionado e convidado a participar da distribuição de carga. Porém, $S_{11}$ rejeita o convite, pois já está envolvido em outra distribuição, iniciada por $S_{12}$. Assim, $S_{11}$ é removido de CANDIDATOS e $S_{10}$ é selecionado porque tem agora a menor carga de trabalho dentre os servidores em CANDIDATOS (Figura \ref{fig:lee2}(c)). Até que a carga de trabalho média dos servidores em SELECIONADOS se torne menor que $0,9 \times CP$, ou seja, $90$, o procedimento acima continua (Figura \ref{fig:lee2}(d) e Figura \ref{fig:lee2}(e)).

@@

Moreover, the upload bandwidth must be taken into account, much more than download. This happens for two reasons: first, the usage of the download bandwidht of each server node grows linearly with the number of players connected to it, while the usage of the upload bandwidth may have a quadratic growth, getting quickly overwhelmed; second, the domestic connections -- which are majority in a volunteer peer-to-peer system -- usually have an upload bandwidth much smaller than the download one.

@@

	\item \textbf{Overhead between two regions}: if there is only a server and a region, comprising the entire virtual environment of the game, the use of upload bandwidth of the server will be proportional to the $W_{total}$. However, due to its distribution among various servers, there is the problem of having players from different regions interacting with each other very close to the border between the regions. Because of that, each state update of these players' avatars will be sent twice. For example, let $A_i$ be the avatar of the player $P_i$, connected to the server $S_i$, and $A_j$ the avatar of the player $P_j$, connected to the server $S_j$. In order for to $P_i$ interacts with $P_j$, it is necessary that $S_i$ send the state of $A_i$ to $S_j$, which then forwards to $P_j$. The same happens in the other way around. The overhead between regions $R_i$ and $R_j$ is equal, therefore, to the sum of interactions between pairs of cells where each one of them is in one of these regions. If $R_i$ and $R_j$ have respectively $m$ and $n$ cells, we have that the interaction -- or \emph{overhead} -- between them is given by:

@@

@@

The choice of cells to include in the region tries to put the most interacting pairs of cells in the same regions, reducing the overhead.% In each step, the algorithm not only tries to get a neighbor cell, but also the one with the highest that not only is free from a nearby cell already present in the region, but whose edge linking them is the heaviest possible. %Figure [TODO: reffig] shows an example of the steps of the growth of a region to achieve its share of the load balancing.

@@

To perform the simulation of the load balancing, a virtual environment was simulated with various avatars in it moving according to the random waypoint model with random wait. However, the choice of which waypoint go was not completely random. Three hotspots were defined, and there was a probability for the havatar choose one of these hotspots as its next waypoint. The purpose of this was to force an uneven distribution of avatars in the virtual environment, putting to test the algorithms of the proposed phase 2. Those who take account of the existence of hotspots form the regions on this basis, reducing the distribution overhead.

@@

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
\bibliography{balancing}   % name your BibTeX data base

>>>>

\newcommand{\ggp}{ProGReGA}
%\newcommand{\ggp}{PGRGA}
\newcommand{\ggpmeaning}{proportional greedy region growing algorithm}
\newcommand{\ggpk}{\mbox{\ggp-KH}}
\newcommand{\ggpkmeaning}{proportional greedy region growing algorithm keeping heaviest cell}
\newcommand{\ggpf}{\ggp-KF}
\newcommand{\ggpfmeaning}{proportional greedy region growing algorithm keeping usage fraction}
\newcommand{\bfa}{BFBCT}
\newcommand{\bfameaning}{best-fit based cell transference}
\newcommand{\kl}{Kernighan\-Lin}
\newcommand{\klalg}{Kernighan-Lin}
\newcommand{\wtodiv}{weight\_to\_divide}
\newcommand{\freecap}{free\_capacity}
\newcommand{\rlist}{region\_list}
\newcommand{\clist}{cell\_list}
\newcommand{\locgroup}{local\_group}
\newcommand{\avguse}{average\_usage}
\newcommand{\wlocal}{local\_weight}
\newcommand{\caplocal}{local\_capacity}
%\newcommand{\wshare}{por\text{\emph{\c{c}}}\tilde{a}o\_da\_carga}
\newcommand{\wshare}{weight\_share}
\newcommand{\wlose}{weight\_to\_lose}
\newcommand{\destreg}{destination\_regions}
\newcommand{\regcap}{free\_capacity}
\newcommand{\wtoreg}{weight\_to\_this\_region}
\newcommand{\misccite}[2]{Available at: #2}
\newcommand{\gamecite}[2]{Available at: #2}

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2006/03/15
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallextended]{svjour3}     % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}         % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage[latin1]{inputenc}
\usepackage{algorithmic}
\usepackage{algorithm}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}
%[TODO:ajustar TODAS as \refs]
\title{A load balancing scheme for MMOGs %\thanks{Grants or other notes
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
}
%\subtitle{Do you have a subtitle?\\ If so, write it here}

@@

Normally, players can freely move their avatars throughout the game world. This makes possible the formation of \emph{hotspots} \cite{ahmed2008mol}, around which the players are more concentrated than in other regions of the virtual environment (Figure \ref{fig:avatarsdistribution}). Moreover, many massively multiplayer online RPGs not only permit but also stimulate, to some extent, the formation of these points of interest. In the worlds of these MMORPGs, there are entire cities, where the players meet to chat, exchange virtual goods or even fight, and there are also desertic areas, with few attractions for the players, and where the number of avatars is relatively small compared to other places of the environment.

\begin{figure}
  \centering
  \includegraphics[width=1.0\linewidth]{images/avatarsdistribution}
  \caption{Distribution of avatars with and without hotspots}
  \label{fig:avatarsdistribution}
\end{figure}

@@

The servers receive the action performed by a player, calculate its outcome and send it to all interested players, who are usually those whose avatars are close to the avatar of the first player. If two players are split between different servers, each of these need not only to send the state update to the player served by it, but it has also to send the update to the server to which the other player is connected. This server, in turn, forwards that state update to the other player (Figure \ref{fig:overheadgeneration}.

\begin{figure}
 \centering
 \includegraphics[width=0.8\textwidth]{images/overheadgeneration}
 \caption{Overhead caused by the interaction of players connected to different servers}
 \label{fig:overheadgeneration}
\end{figure}

@@

@@

%TODO: explicar porque cresce linearmente
Moreover, the upload bandwidth must be taken into account, much more than download. This happens for two reasons: first, the usage of the download bandwidht of each server node grows linearly with the number of players connected to it, while the usage of the upload bandwidth may have a quadratic growth, getting quickly overwhelmed; second, the domestic connections -- which are majority in a volunteer peer-to-peer system -- usually have an upload bandwidth much smaller than the download one.

%TODO: translate
Há questões de menor importância, mas que também devem ser levadas em consideração e que não foram consideradas em outros trabalhos. Uma delas é que dificilmente o sistema servidor será homogêneo -- considerando-se que é baseado em recursos voluntários. Portanto, não se pode assumir que os servidores tenham a mesma quantidade de recursos. Outro problema, específico do algoritmo proposto por \cite{lee2003sdl}, é que o critério de alocar para cada servidor uma carga equivalente a 90\% da sua capacidade, com o fim de evitar rebalanceamentos constantes, é fraco. Não há garantias de que o sistema servidor terá 11,11\% a mais de capacidade total do que o necessário. Muito pelo contrário, ele deve se adaptar a situações de sobrecarga generalizada. Além disso, o critério de parada é fraco pois o algoritmo não termina se o sistema todo estiver com carga acima de 90\% da sua capacidade, ou se o servidor sobrecarregado não tiver vizinhos disponíveis para balanceamento.

There are also other issues which must not be overlooked. One of them is that the server system is probably not homogenous -- considering that it is based on volunteer resources. Therefore, one can not assume that the servers have the same amount of cpu power or network bandwidth. Another problem, specific to the algorithm proposed by \cite{lee2003sdl}, is that the criterion to allocate to each server a load equivalent to 90\% of its capacity in order to avoid constant rebalancing is weak. There are no guarantees that the server system has 11.11\% more total capacity than the necessary. On the contrary, it should adapt to situations of widespread overload. Moreover, the criterion for stopping is weak because the algorithm does not stop when the whole system is loaded above 90\% of its capacity, or if the overloaded server has no neighbors capable of receiving more load.

%TODO: translate
Porém, uma idéia importante que foi sugerida é a do uso de grafos para representar o ambiente virtual e de usar algoritmos de particionamento de grafo para realizar o balanceamento de carga. No capítulo \ref{cap:contexto}, foi feita uma introdução a respeito desse princípio, que é a base do esquema de balanceamento de carga aqui proposto. Na seção seguinte, será mostrado como um ambiente virtual de um MMOG pode ser mapeado em um grafo.

@@

	\item \textbf{Overhead between two regions}: if there is only a server and a region, comprising the entire virtual environment of the game, the use of upload bandwidth of the server will be proportional to the $W_{total}$. However, due to its distribution among various servers, there is the problem of having players from different regions interacting with each other very close to the border between the regions (Figure \ref{fig:interactingregions}). Because of that, each state update of these players' avatars will be sent twice. For example, let $A_i$ be the avatar of the player $P_i$, connected to the server $S_i$, and $A_j$ the avatar of the player $P_j$, connected to the server $S_j$. In order for to $P_i$ interacts with $P_j$, it is necessary that $S_i$ send the state of $A_i$ to $S_j$, which then forwards to $P_j$. The same happens in the other way around. The overhead between regions $R_i$ and $R_j$ is equal, therefore, to the sum of interactions between pairs of cells where each one of them is in one of these regions. If $R_i$ and $R_j$ have respectively $m$ and $n$ cells, we have that the interaction -- or \emph{overhead} -- between them is given by:

@@

	\begin{figure}
	\centering
	\includegraphics[width=0.8\textwidth]{images/interactingregions}
	\caption{Players interacting across a border between regions}
	\label{fig:interactingregions}
	\end{figure}

@@

%The choice of cells to include in the region tries to put the most interacting pairs of cells in the same regions, reducing the overhead.% In each step, the algorithm not only tries to get a neighbor cell, but also the one with the highest that not only is free from a nearby cell already present in the region, but whose edge linking them is the heaviest possible. %Figure [TODO: reffig] shows an example of the steps of the growth of a region to achieve its share of the load balancing.

%A escolha de células para incluir na região busca fazer com que cada uma das arestas mais pesadas do grafo $GW$ ligue vértices da mesma partição, reduzindo o corte de aresta e, assim, o overhead. A cada passo é dada preferência a escolher a célula livre que não apenas seja vizinha de uma célula já presente na região, mas também cuja aresta ligando-as seja a mais pesada possível. A Figura \ref{fig:ggp} mostra um exemplo dos passos do crescimento de uma região até atingir a sua parcela de carga no balanceamento.

The criterion to choose a cell to include in the region is to make the heaviest edges on the graph $GW$ connect vertices in the same partition, reducing the edge-cut and, thus, the overhead. In each step, it is selected the cell which not only is adjacent to a cell already present in the region, but whose edge connecting them is the heaviest possible. Figure \ref{fig:ggp} illustrates the steps of growth of a region until it reaches its load share.

%TODO: referenciar karypis e kumar, wow e outros jogos (eve-online, warhammer e cabal)
\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{images/ggp}
\caption{Growth of a partition (region) with \ggp}
\label{fig:ggp}
\end{figure}

No exemplo da figura \ref{fig:ggp}, há dois servidores, $S_1$ e $S_2$, sendo que $p(S_1) = 30$ e $p(S_2) = 18$. A carga total do ambiente sendo reparticionado é $W_{total} = 32$. Para que a divisão seja proporcional à capacidade dos servidores, as parcelas de carga designadas a $S_1$ e $S_2$ são de 20 e 12, respectivamente. A seleção começa pelo vértice de peso 6 (célula livre com maior carga) e a partir daí, é adicionado a cada passo o vértice ligado pela aresta mais pesada. As arestas selecionadas e os vértices que pertencem à nova partição estão destacados.

@@

To perform the simulation of the load balancing, a virtual environment was simulated with various avatars in it moving according to the random waypoint model \cite{bettstetter2004spr}. However, the choice of which waypoint go was not completely random. Three hotspots were defined, and there was a probability for the havatar choose one of these hotspots as its next waypoint. The purpose of this was to force an uneven distribution of avatars in the virtual environment, putting to test the algorithms of the proposed phase 2. Those who take account of the existence of hotspots form the regions on this basis, reducing the distribution overhead.

@@

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
\bibliography{dissertation}   % name your BibTeX data base
